Manifold Destiny - or how to rig an election

### 2024-01-07 paper.qmd
moreupd

```{r, echo=F, message=F, warning=F}
library(ManifoldDestiny)
ManifoldDestiny::wasmnonload()
metad <- jsonlite::fromJSON(paste0(rprojroot::find_rstudio_root_file(),"/data-raw/metadata.json"))
md <- jsonlite::fromJSON(paste0(rprojroot::find_rstudio_root_file(),"/data-raw/metadata.json"))
source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/simulations.R'))
source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/applications.R'))
modlat <- eqpar$meql
```

# `r #metad$toc$sec1$nmn`

@@

**Our numerical example**

:::: {.columns}
::: {.column width="50%"}
Case 1
:::
```{r, echo=F}
clr <- c("black","green","red","blue")
lego_hux_imp_i <- as_hux(towi[,1:3]) %>%
  set_text_color(5, 1:2, clr[1]) %>%
  set_background_color(1, 1:3, clr[2]) %>% 
  set_background_color(2, 1:3, clr[3]) %>%
  set_background_color(3:5, 1:3, clr[4]) 
lego_hux_imp_i
```

::: {.column width="50%"}
Case 2
:::
```{r, echo=F}
lego_hux_imp_ii <- as_hux(towii[,1:3]) %>%
  set_text_color(5, 1:2, clr[1]) %>%
  set_background_color(1, 1:3, clr[2]) %>% 
  set_background_color(2, 1:3, clr[3]) %>%
  set_background_color(3:5, 1:3, clr[4]) 
lego_hux_imp_ii
```
::::

@@

Numerical calculation of candidate one total percentage
- Precinct 1
 - $\alpha=\frac{`r cg[1]`+`r ch[1]`\cdot `r cGamma[1]`}{1+`r cGamma[1]`}=`r calpha[1]`$
- Precinct 2
 - $\alpha=\frac{`r cg[2]`+`r ch[2]`\cdot `r cGamma[2]`}{1+`r cGamma[2]`}=`r calpha[2]`$
- Precinct 3
 - $\alpha=\frac{`r cg[3]`+`r ch[3]`\cdot `r cGamma[3]`}{1+`r cGamma[3]`}=`r calpha[3]`$

@@

@@

[Link to R2 simulator for testing](https://shinylive.io/r/app/#h=0&code=NobwRAdghgtgpmAXGKAHVA6ASmANGAYwHsIAXOMpMAdzgCMAnRRASwgGdSoAbbgCgA6YALJQILAGZFuAEwAicTmwCeAdQCCAZWEAFIbgAEDOKiLsDAXgNCAFqVKp2iAPTPuRLgxZEb1AEwAbBgA5iykNgCudBjezsJyAPoa2gDCQgCUAhC0jMxsnDz8QqLiUrIKShBqWvH6RiZmltZgdg5Oru6e3r6BIWGR0bHxSTVpYJkQ3CyMUAzKfCWS0vKKpCrJuhNTM3MLYkvlq+s1clvTDLPz7DYqZzvzMFDBXvYMdxe7wcGonX4TRKg1iR2Hx2AQWKgKBYAJyw96XPg-DzcZTw3ZyAAqaPmAGsoHRuHAAKIAD1IF2xfBkP2Ubyy2w+8zsMG4pCI0nYlOeLBkpPJUAmERYBgAPABaAwSbhCmQ6J5wQQQAwGNakQlyiBwIpgLB+AyaFgwCLcKBshgZXBZZVcOjsOCkDVaxXK634x3agDiF1QNjqVpdAcD7B59FmABkoMoiBFSM7A-GXcGZKGGO64wmM3bCQRSABJCCoGOCMAAN1mLHxhL9YE0WrgOYMUElRAYMEQdQIxYAci3Htxm62hE0hBBezxqwAJZSMHkDmBDqy2adeGTVgDy6DMYW8Sqkg7Aw7AANMwaBEAy6UtSozGYgEXgXgI+cLsZHqGMBGrXfvdDgDAMRASAY771mwOZOHUZbSnATQAMwAAzwZe-o3vG5Bks+RZCMYwQltWWBwKEnB-nAMgGCWHh-uYfDwGIuCcDImR4ORPARHAFgAOSwRgACMACshjwRgfh8RxyHXqhgboXmBZYWAMBsI8JLVsIbCNhAZFKSx0HmHeMC-v+gHkZRDDsJBrHsRxAAsiGGDZiFiVekloXAGGya+YCkBEDDnsxQg6AwRB0PiLBTKQygAUBFFrBAwTtsxUFsU0HFCQJBhCTx8GOShznWq5MkvsWbAlpC5p+WARLZmeBjsMoxEwCxXghWFyjxYYiUwZxqWCbxWXiblAbSZhHkRNWAVBc1YQRUZJLOMEzgQG12lJVYKUYAEPWZdlEkDUN7nFjIeHlXIkhAe+E10KFU2RQYs3zYt5nQclYpCX4PXwQAHNtA0unpf4sE++1CDIFzUGZ5Xfvpf43SDUBg49K0GAh-U-Y2OY7gAQjGbIQMWDARL5hhCFgBP6oaxqmjuGQ5ZJKO5Y8bBpjTzlIqQa4xoVQisxazOSTahLsw4cn83A1M7bTvPxukEwBnTKpumITpCBilaKH64sZjadoOorRQazeNpM2AKsEjBPH6JLqNJimEZRlh+uo4mIbBamut8HLjuBgzEBMw7nuuqbguc55qvm+MluOzL-uyxHgYe6hhtu8rqsGH4Ft+z91su7b0avhnVvO7Mabx9H3u+9Hmuq0Hwuq2n4f5z9UcVwYTfOa3N6txMEwAMTdwYCgSGwMF2gwJZ-lkI9j-+4qSgT6MkHwbAvoYucvukBggP6-p9yTSrBkaJpnv6MgSN8ooSnAY9kARUDo2Pi-uQAJPjECGJvO198N-cAGLmztOElnPo2dgGA-qPlBOSdgPwwgPxfM-QiJYiZ4AyMAYAPEAC66DW4KQgFpGeUAQFgIBhA0y0DYxLxjI-HBSkkGWnGKgjBWCcpeR8kAghoD7z-U7JwUhYVYGUJYa-ZodD0gMMwa3YqpU2GEM4eAnhUC+EUNII-SRf5aEoLQeInKERpEcIfMQ+RZD+HKIiOo+hmimE7UOrooh3DIFGKUY-Q6ZjREWNbqQICM9jDQIIKaBUjjYZg1wAwPw+9cYFg-E0RxET6xORvDhcwVhOwAM0YYFJfgsFxIzNQqAJImidhySSVJhTgAZMvCqbySokl8EEakipPlSmZIMPmEq0NqmqIYHUjpjSS4uh0dUiIdTBllKyQmax1TDp1MmSMiOe4GpWB4qIzsPFcBvVgukSxAYT5nxnjIU0UAMASAuPAPgISoAWAJlMTgNSJDADtAARz4CswksVwg3MvOszJIS6AXMmCwa5Hi7lwEeW9F5wQ3keI+Rs6WBgACkAA+WFBgjRcHIIiP8BAmhkFCgqM5hhMpIXXjlT+7kf5px2qVAgFA1iEnMDPTs0J4KGGhGlOEOUOCkDIvSp5qzcB8VbgxIBtFcbbNQM-PwAoADUHKZAACp2AyCpKfMVZzW6Up4kA+5EQxA0oVKK8VUBcDnVtE0Sl1KcXmGcAYAlaqMV6hnlqnVOKlXfHFXQI1gUTVWDNdi2lBgrU2pypSk+QC9lcEOcchU6qjV2twDK+inL14IqRWGg5ItmDa2gPAEEnZ-IYqMBKi2YAdD5u+UWrsDEi0ADVK3jCJTtK5sZRUxoYAQE+UcAC+Uc+4pGMH44CnR-Sr0oazIBxgNJ-h0J0Pg78BCkGVN3RgRAwZ-ndjlE+qBQ3KvdmI9BcLEUqh5LSZgqAWDRQSO4WKq7iDcHYBYXNYAzlFrLXW-dKaaRMEQLMC4V6+BZrgC2ggrcvisyVagXAUBFB8BJBYDqhhB68AsP+mFkrrAa2CHAIgMAEg3E4EQZ4sBERbmqouMAIZsXhTqDwH0TYrDdQMJdDgcEkIGFQ8zE0toalhEJIeCc-y2QEYakZKtFlzB0AiikPxwQWzKDqHk0jInoJ1AiqRlI0YyAZFY8zDDWGEglimJqaDbByCtpMPOqw7DbEutQDuxp6C0G4AweUgz9plCQkPHs64pEOzSBbIeAkbFNOSu05h7D+mh5GbIBiszTRLOyOIaK2zGT7NvSc4YFz4V3Okc8zYbzzEb1+dIwF0W4wtPodC3plzkWTNUsBLFmR+jOyJdccl4AsFHNNIy25zqzQct5cMAV-8RWlOleC+V3T4XDMkmM9FurFmGtcOs0lzBbXcBWU60PTLPWhDsGkDyHz7ghvNGMKuUbIXdMcadMoCwTKYNxca0tlrK2Vlpcu9wCwQhZX6B0zAD7YA3tBfO9ht7fBru3YsPdxbzXUGtdS5kt7f3ZVfbwD9v7AOzvjeB-iK7N3cB3YW+A6HdnVuvex+9z7SPvuhbR2TwHGtwhwHgAkHBhoeDuzK-GMEPA4AJAQ9wZnYhtX8A6okgwD7u6IQAOx+AxnXJB3ciQBGhN-RCF4F01XtJKUK-ZBtmWvF2re15e4GF7XAftIsh0cwEaHMdFBkwMAAKq5hnTlTEzBU0i1OepxVRPku4BkCwUIpA71WWlv6A315h3KJFvaiU477dO5dztN3iAPeqy9wTH327XGMMMAHoPIfykAjPKLxtiJ5RhgoOCmwTQ-CEvDxMDtWQd6kzQKgCeNwqjqHQHwIUTQhSGEnm0jXo8-wTCyGADt6CgA)

@@

@@

[@delfino2011analysis]

@@

@@

@@

@@

@@

@@

>>>>

```{r, echo=F, message=F, warning=F, results='hide'}
#Sys.info()[['sysname']]=="Emscripten"
ManifoldDestiny::wasmconload()
metad <- jsonlite::fromJSON(paste0(rprojroot::find_rstudio_root_file(),"/data-raw/metadata.json"))
md <- jsonlite::fromJSON(paste0(rprojroot::find_rstudio_root_file(),"/data-raw/metadata.json"))
source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/20laws40isos.R'))
source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/simulations.R'))
source(paste0(rprojroot::find_rstudio_root_file(),'/script/R/applications.R'))
modlat <- eqpar$meql
library(huxtable)
library(kableExtra)
```

# `r metad$toc$sec1$nmn`

@@

**Our two stylized numerical example**

:::: {.columns}
::: {.column width="50%"}
Case 1
:::
```{r, echo=F}
clr <- c("black","green","red","blue")
lego_hux_imp_i <- as_hux(towi[,1:3]) %>%
  set_text_color(5, 1:2, clr[1]) %>%
  set_background_color(1, 1:3, clr[2]) %>% 
  set_background_color(2, 1:3, clr[3]) %>%
  set_background_color(3, 1:3, clr[4]) %>%
  set_background_color(4, 2:3, clr[4]) %>%
  set_background_color(5, 3:3, clr[4]) %>%
  set_font_size(16)

lego_hux_imp_i
```

::: {.column width="50%"}
Case 2
:::
```{r, echo=F}
lego_hux_imp_ii <- as_hux(towii[,1:3]) %>%
  set_text_color(5, 1:2, clr[1]) %>%
  set_background_color(1, 1:3, clr[2]) %>% 
  set_background_color(2, 1:3, clr[3]) %>%
  set_background_color(3, 1:3, clr[4]) %>%
  set_background_color(4, 2:3, clr[4]) %>%
  set_background_color(5, 3:3, clr[4]) %>%
  set_font_size(16)

lego_hux_imp_ii
```
::::

@@

Numerical calculation of candidate one total percentage

- Precinct 1
  - $\alpha=\frac{`r cg[1]`+`r ch[1]`\cdot `r cGamma[1]`}{1+`r cGamma[1]`}=`r calpha[1]`$
- Precinct 2
  - $\alpha=\frac{`r cg[2]`+`r ch[2]`\cdot `r cGamma[2]`}{1+`r cGamma[2]`}=`r calpha[2]`$
- Precinct 3
  - $\alpha=\frac{`r cg[3]`+`r ch[3]`\cdot `r cGamma[3]`}{1+`r cGamma[3]`}=`r calpha[3]`$

---

@@

[Application 1: Election process simulator]({{< var shn1 >}})

---

@@

@@

```{r}
#kable(laws_df1, format = "markdown", escape = FALSE) %>% kableExtra::kable_paper(full_width = F) %>% kableExtra::kable_styling(font_size=12)
```

@@

@@

:::: {.columns}

::: {.column width="50%"}

*Standard form*

Setting the first three parameters:

- $x_i=\frac{s_i}{s_i+t_i}$
- $y_i=\frac{u_i}{u_i+v_i}$
- $\zeta_i=\frac{u_i+v_i}{s_i+t_i}$

Backsolve for

- $\alpha_i(x_,y_i,\zeta_i)=\frac{x_i+\zeta_i\cdot y_i}{\zeta_i+1}$
- $\lambda_i(x_i,y_i,\zeta_i)=\frac{x_i\zeta_i\cdot(1-y_i)}{\zeta_i+1}$

::: {.column width="50%"}

*Opposition form*

Setting the first three parameters:

- $g_i=\frac{s_i}{s_i+t_i}$
- $h_i=\frac{u_i}{u_i+v_i}$
- $\Gamma_i=\frac{u_i+d_i}{u_i+v_i}$

Backsolve for

- $\alpha_i(g_i,h_i,\Gamma_i)=\frac{x_i+\zeta_i\cdot y_i}{\zeta_i+1}$
- $\Omega_i(g_i,h_i,\Gamma_i)=\frac{x_i\zeta_i\cdot(1-y_i)}{\zeta_i+1}$

::::

---

**Properties**:

- 1<sup>st</sup> The total election outcome will be determined by *three* parameters: ( $x_i, y_i,\zeta_i$ ) under standard form and ( $g_i, h_i, \Gamma_i$ ) using hybrid form. 
- 2<sup>nd</sup> Changes in $\Gamma_i$ and $\zeta_i$ would in general have an impact on $\alpha_i$.

@@

**Setting the prefered election outcome**

:::: {.columns}

::: {.column width="40%"}

Non-rigged election using standard form
```{r, echo=F, echo=F, out.width="150%", fig.cap=""}
#0app_n_out[[4]] 
```

::: {.column width="60%"}

Rigged election using standard form
```{r, echo=F, echo=F, out.width="150%", fig.cap=""}
#app
```

::::

---

**Normal**
<br>
\begin{equation*}
f(x)_{0}^{1}=
`r #paste0(round(polynom::polynomial(unname(pri_int_ex1)),3))`$
\end{equation*}

Election outome
```{r, echo=F}
#sum(dplyr::select(app_n_out[['rdfc']],S,U))/sum(dplyr::select(app_n_out[['rdfc']],S,T,U,V))
```

<br>
**Rigged**
<br>

$f(x)_{0}^{1}=`r #paste0(round(polynom::polynomial(unname(pos_int_ex1)),3))`$

Election outome
```{r, echo=F}
#sum(dplyr::select(app_ex1_out[['rdfc']],S,U))/sum(dplyr::select(app_ex1_out[['rdfc']],S,T,U,V))
```

---

### Rigged election example 1: applying normal form

- Setting for eaching precinct the first three parameters
 - $\alpha_i=<...>$ (i.e., equal to the prefered election outcome)
 - $x_i=\frac{a_i}{a_i+b_i}$ (true observations)
 - $y_i=\overset{\text{Manifold equation}}{\overbrace{\overset{}{M(x_i,\alpha_i,\underset{\text{outside information}}{...})}}}+\epsilon_t \text{ where } \epsilon_t \sim N(0,\sigma^2)$
- Backsolve for
 - $\zeta_i(x_i,\alpha_i,y_i)=\frac{x_i-\alpha_i}{\alpha_i-y_i}$
 - $\lambda_i(x_i,\zeta_i,y_i)=\frac{x_i-\zeta_i (1-y_i)}{\zeta_i+1}$

---

**Properties**:

- 3<sup>rd</sup> $y_i$ can now be determined by *two* equations
  - a. Tautology (must hold)
  - b. Manifold equation + error term (white noise)
- 4<sup>th</sup>. Changes in $\zeta_i$ - and as a corrollary $\Omega_i$ - is now 
  - a. Solely caused by the error term $\epsilon_i$ and, if used, the non-linear part of the Manifold equation.
  - b. Has no longer any causal impact on $\alpha_i$

---

### Rigged election example 2: applying hybrid form

- Setting for eaching precinct the first three parameters
 - $\alpha_i=<...>$ (i.e., equal to the prefered election outcome)
 - $h_i=\frac{s_i}{s_i+v_i}$ (true observations)
 - $g_i=\overset{\text{Manifold equation}}{\overbrace{\overset{}{M(h_i,\alpha_i,\underset{\text{outside information}}{...})}}}+\epsilon_t \text{ where } \epsilon_t \sim N(0,\sigma^2)$
- Backsolve form
 - $\Gamma_i(g_i,\alpha_i,h_i)=\frac{g_i-\alpha_i}{\alpha_i-h_i}$
 - $\Omega_i(g_i,\Gamma,h_i)=-\alpha_i+\frac{\Gamma_i+2g_i}{\Gamma_i+1}$

---

**Properties**:

- 5<sup>th</sup> $y_i$ can now be determined by *two* equations
  - a. Tautology (must hold)
  - b. Manifold equation + error term (white noise)
- 6<sup>th</sup> Changes in $\Gamma_i$ - and as a corrollary $\lambda_i$ - is now 
  - a. Solely caused by the error term $\epsilon_i$ and, if used, the non-linear part of the Manifold equation.
  - b. Has no longer any causal impact on $\alpha_i$

---

### Rigged election example 3: applying opposition form

- Setting for eaching precinct the first three parameters
 - $\lambda_i=<...>$ (i.e., equal or less than the prefered election outcome)
 - $n_i=\frac{a_i}{a_i+c_i}$ (true observations)
 - $m_i=\overset{\text{Manifold equation}}{\overbrace{\overset{}{M(n_i,\lambda_i,\underset{\text{outside information}}{...})}}}+\epsilon_t \text{ where } \epsilon_{t_i} \sim N(0,\sigma^2)$
- Backsolve for
 - $\Omega_i(n_i,m_i,\lambda_i) = \frac{\lambda_i(n_i-m_i)-m_i(2n_i-1)}{1-(n_i+m_i)}$
 - $\xi_i(n_i,m_i,\Omega_i)=\frac{m_i-\Omega_i}{\Omega_i-n_i}$

---

**Properties**:

- 7<sup>th</sup> $m_i$ can now be determined by *two* equations
  - a. Tautology (must hold)
  - b. Manifold equation + error term (white noise)
- 8<sup>th</sup> Changes in $\xi_i$ is now 
  - a. Solely caused by the error term $\epsilon_i$ and, if used, the non-linear part of the Manifold equation.
  - b. Has no longer any causal impact on $\lambda_i$

---

**Setting the prefered election outcome**

:::: {.columns}

::: {.column width="50%"}

Non-rigged election using standard form
```{r, echo=F, echo=F, out.width="150%", fig.cap=""}
app_n_out[[4]] 
```

::: {.column width="50%"}

Rigged election using standard form
```{r, echo=F, echo=F, out.width="150%", fig.cap=""}
app_ex1_out[[4]] 
```
:::

::::

---

**Normal**
<br>
\begin{equation*}
f(x)_{0}^{1}=
`r #paste0(round(polynom::polynomial(unname(pri_int_ex1)),3))`$
\end{equation*}

Election outome
```{r, echo=F}
sum(dplyr::select(app_n_out[['rdfc']],S,U))/sum(dplyr::select(app_n_out[['rdfc']],S,T,U,V))
```

<br>
**Rigged**
<br>

$f(x)_{0}^{1}=`r #paste0(round(polynom::polynomial(unname(pos_int_ex1)),3))`$

Election outome
```{r, echo=F}
sum(dplyr::select(app_ex1_out[['rdfc']],S,U))/sum(dplyr::select(app_ex1_out[['rdfc']],S,T,U,V))
```

---

@@

- Loss function
  - $M(,)=$x+y$
    - Use: Linear, quadratic, cubic or quartic (all have analytical solutions)
    - Also: Euler rotation
  - System approach Demands mulitvariate Manifold Equations
    - Abc
    - Def
  - Constrain on implementation
    - Flipping votes on one mode
    - Vote registration until upper limit

---

```{r, echo=F, out.width="80%", fig.cap=""}
bI <- paste0(rprojroot::find_rstudio_root_file(),'pngs/loss_5_conditions.png')
knitr::include_graphics(bI)
```

@@

---

@@

```{r, echo=FALSE, results='asis',warning=F}
params <- list(name="2")
app1_fil <- paste0(rprojroot::find_rstudio_root_file(), '/app_par.qmd')
text <- readLines(app1_fil)
cat(knitr::knit_child(text = paste(text[50:length(text)], collapse = "\n"), quiet = TRUE))
```
---

[xxx application]({{< var shn3 >}})

---

@@

[xxx application]({{< var shn4 >}})

---
